---
title: "Hands-on_Ex04a"
author: "Huang Zhanyi"
date: "May 5, 2026"
date-modified: "last-modified"
execute: 
  eval: true
  echo: true
  warning: false
  freeze: true
---

# Visualising Distribution

## Install packages

```{r}
pacman::p_load(ggdist, ggridges, ggthemes,
               colorspace, tidyverse)
```

## Import Data

```{r}
exam_data <- read_csv("/Users/zhanyi/Desktop/M Y1S2/ISSS608/ISSS608-VAA/project 1/ISSS608/Hands-on_Ex/Hands-on_Ex04a/data/Exam_data.csv")
```

```{r}
exam_data$MATHS <- as.numeric(exam_data$MATHS)

```

```{r}
exam_data$ENGLISH <- as.numeric(exam_data$ENGLISH)

```

```{r}
exam_data$SCIENCE <- as.numeric(exam_data$SCIENCE)
```

## Visualising Distribution with Ridgeline Plot

### Plotting ridgeline graph: ggridges method

```{r}
ggplot(exam_data, 
       aes(x = ENGLISH, 
           y = CLASS)) +
  geom_density_ridges(
    scale = 3,
    rel_min_height = 0.01,
    bandwidth = 3.4,
    fill = lighten("#7097BB", .3),
    color = "white"
  ) +
  scale_x_continuous(
    name = "English grades",
    expand = c(0, 0)
    ) +
  scale_y_discrete(name = NULL, expand = expansion(add = c(0.2, 2.6))) +
  theme_ridges()
```

### Varying fill colors along the x axis

```{r}
ggplot(exam_data, 
       aes(x = ENGLISH, 
           y = CLASS,
           fill = stat(x))) +
  geom_density_ridges_gradient(
    scale = 3,
    rel_min_height = 0.01) +
  scale_fill_viridis_c(name = "Temp. [F]",
                       option = "C") +
  scale_x_continuous(
    name = "English grades",
    expand = c(0, 0)
  ) +
  scale_y_discrete(name = NULL, expand = expansion(add = c(0.2, 2.6))) +
  theme_ridges()
```

### Mapping the probabilities directly onto colour

```{r}
ggplot(exam_data,
       aes(x = ENGLISH, 
           y = CLASS, 
           fill = 0.5 - abs(0.5-stat(ecdf)))) +
  stat_density_ridges(geom = "density_ridges_gradient", 
                      calc_ecdf = TRUE) +
  scale_fill_viridis_c(name = "Tail probability",
                       direction = -1) +
  theme_ridges()
```

### Ridgeline plots with quantile lines

```{r}
ggplot(exam_data,
       aes(x = ENGLISH, 
           y = CLASS, 
           fill = factor(stat(quantile))
           )) +
  stat_density_ridges(
    geom = "density_ridges_gradient",
    calc_ecdf = TRUE, 
    quantiles = 4,
    quantile_lines = TRUE) +
  scale_fill_viridis_d(name = "Quartiles") +
  theme_ridges()
```

```{r}
ggplot(exam_data,
       aes(x = ENGLISH, 
           y = CLASS, 
           fill = factor(stat(quantile))
           )) +
  stat_density_ridges(
    geom = "density_ridges_gradient",
    calc_ecdf = TRUE, 
    quantiles = c(0.025, 0.975)
    ) +
  scale_fill_manual(
    name = "Probability",
    values = c("#FF0000A0", "#A0A0A0A0", "#0000FFA0"),
    labels = c("(0, 0.025]", "(0.025, 0.975]", "(0.975, 1]")
  ) +
  theme_ridges()
```

## Visualising Distribution with Raincloud Plot

### Plotting a Half Eye graph

```{r}
ggplot(exam_data, 
       aes(x = RACE, 
           y = ENGLISH)) +
  stat_halfeye(adjust = 0.5,
               justification = -0.2,
               .width = 0,
               point_colour = NA)
```

### Adding the boxplot with geom_boxplot()

```{r}
ggplot(exam_data, 
       aes(x = RACE, 
           y = ENGLISH)) +
  stat_halfeye(adjust = 0.5,
               justification = -0.2,
               .width = 0,
               point_colour = NA) +
  geom_boxplot(width = .20,
               outlier.shape = NA)
```

### Adding the Dot Plots with stat_dots()

```{r}
ggplot(exam_data, 
       aes(x = RACE, 
           y = ENGLISH)) +
  stat_halfeye(adjust = 0.5,
               justification = -0.2,
               .width = 0,
               point_colour = NA) +
  geom_boxplot(width = .20,
               outlier.shape = NA) +
  stat_dots(side = "left", 
            justification = 1.2, 
            binwidth = .5,
            dotsize = 2)
```

### Finishing touch

\`\`\`{r} ggplot(exam_data, aes(x = RACE, y = ENGLISH)) + stat_halfeye(adjust = 0.5, justification = -0.2, .width = 0, point_colour = NA) + geom_boxplot(width = .20, outlier.shape = NA) + stat_dots(side = "left", justification = 1.2, binwidth = .5, dotsize = 1.5) + coord_flip() + theme_economist() ---
