---
title: "Visual Analysis"
subtitle: "Exploring Profitability and Volatility — Spanish Motor Insurance Portfolio"
author: "Huang Zhanyi"
date: "May 16, 2026"
date-modified: "last-modified"
execute: 
  eval: true
  echo: true
  warning: false
  freeze: true
---

# Overview

This report investigates the **profitability** (loss ratio = total incurred / total premium) and **volatility** (variability of the loss ratio) of a Spanish motor insurance portfolio containing 354,140 policy-year records across 2022 to 2024. The analysis proceeds in four stages: data wrangling, univariate exploration of key variables, bivariate analysis to identify drivers of loss ratio, and interactive visualisations for deeper exploration.

------------------------------------------------------------------------

# 1. Data Wrangling

The raw file uses a **semicolon** as the column separator and a **dot** as the decimal mark. `read_delim()` is used with explicit `delim` and `locale` arguments to handle this correctly. Using `read_csv()` would read all 47 columns as a single text column and fail.

```{r}
#| label: setup
pacman::p_load(tidyverse,
               ggrepel, ggthemes, hrbrthemes, patchwork,
               ggiraph, plotly, DT,
               ggridges, ggdist, ggstatsplot, colorspace,
               FunnelPlotR, knitr, scales)
set.seed(1234)
```

## 1.1 Load the Data

```{r}
#| label: load-raw
policy_raw <- read_delim(
  "data/Dataset of motor insurance portfolio.csv",
  delim          = ";",
  locale         = locale(decimal_mark = "."),
  show_col_types = FALSE
)

cat("Rows:   ", nrow(policy_raw), "\n")
cat("Columns:", ncol(policy_raw), "\n")
```

The raw dataset contains **354,140 rows and 47 columns**. Each row is one policy-year record. The same policy can appear across multiple years if renewed.

------------------------------------------------------------------------

## 1.2 Select Relevant Columns

The 47 columns include coverage-level breakdowns of premium, claims and incurred amounts (e.g. `liability_premium`, `theft_claims`, `fire_incurred`). These are sub-components of the three totals — `total_premium`, `total_claims` and `total_incurred` — which are already retained. Since the objective is **portfolio-level** profitability and volatility, the breakdowns add no analytical value. `select()` retains only the 19 columns needed.

```{r}
#| label: select-cols
policy <- policy_raw %>%
  select(
    insured_id, year,
    policy_type, policy_status, business_type,
    payment_frequency, bonus_score,
    driver_age, vehicle_age, age_driving_licence,
    fuel_type, vehicle_value, vehicle_brand,
    municipality_type, circulation_area,
    total_premium, total_incurred,
    total_claims, total_exposure
  )

cat("Columns kept:   ", ncol(policy), "\n")
cat("Columns removed:", ncol(policy_raw) - ncol(policy), "\n")
```

The dataset is reduced from **47 to 19 columns**. The 28 removed columns are all coverage-level breakdowns (liability, property, theft, fire, glass, legal protection, occupants) for premium, claims and incurred amounts.

------------------------------------------------------------------------

## 1.3 Check for Missing Values

Missing values are identified before any transformation. `summarise(across(everything(), ~ sum(is.na(.))))` counts `NA` values across all columns simultaneously.

```{r}
#| label: missing-check
policy %>%
  summarise(across(everything(), ~ sum(is.na(.)))) %>%
  pivot_longer(everything(),
               names_to  = "column",
               values_to = "missing_count") %>%
  filter(missing_count > 0) %>%
  mutate(missing_pct = round(missing_count / nrow(policy) * 100, 2)) %>%
  knitr::kable(caption = "Columns with missing values")
```

Only four columns have missing values. The two largest are `fuel_type` (1,287 rows, 0.36%) and `vehicle_value` (513 rows, 0.14%). Both are negligible. No imputation is applied — missing rows are excluded at plot time using `filter(!is.na(...))`.

------------------------------------------------------------------------

## 1.4 Cast Factors and Derive New Variables

Two transformations are applied:

**Factors** — eight categorical columns are cast to ordered factors with readable labels. This controls axis ordering in plots and replaces raw codes (e.g. "A", "G") with meaningful labels (e.g. "Active", "Good"). `vehicle_brand` is simplified to the top 10 brands with all others collapsed to "Other".

**Derived variables** — seven new columns are computed. These are the core metrics used in every subsequent visual analysis.

```{r}
#| label: transform
policy <- policy %>%
  mutate(
    year = as.integer(year),

    policy_type = factor(policy_type,
                         levels = c("TP","TPG","CC","COMP_E","COMP_N")),

    policy_status = factor(policy_status,
                           levels = c("A","C"),
                           labels = c("Active","Cancelled")),

    business_type = factor(business_type,
                           levels = c("NB","P"),
                           labels = c("New Business","Portfolio")),

    payment_frequency = factor(payment_frequency,
                               levels = c("A","S","Q"),
                               labels = c("Annual","Semi-annual","Quarterly")),

    bonus_score = factor(bonus_score,
                         levels = c("G","N","B"),
                         labels = c("Good","Neutral","Bad")),

    fuel_type = factor(fuel_type,
                       levels = c("D","G"),
                       labels = c("Diesel","Gasoline")),

    municipality_type = factor(municipality_type,
                               levels = c("I","C","IS"),
                               labels = c("Inland","Coastal","Islands")),

    circulation_area = factor(circulation_area,
                              levels = c("U","R"),
                              labels = c("Urban","Rural")),

    vehicle_brand = if_else(
      vehicle_brand %in% (count(., vehicle_brand, sort = TRUE) %>%
                          slice_head(n = 10) %>% pull(vehicle_brand)),
      vehicle_brand, "Other"),
    vehicle_brand = factor(vehicle_brand),

    loss_ratio      = total_incurred / total_premium,
    claim_frequency = total_claims   / total_exposure,
    severity        = total_incurred / total_claims,
    has_claim       = total_claims   > 0,

    loss_ratio_bins = cut(loss_ratio,
                          breaks = c(-Inf, 0, 0.5, 1, Inf),
                          labels = c("Zero","0-0.5","0.5-1","Above 1")),

    driver_age_band = cut(driver_age,
                          breaks = c(17,25,35,45,55,65,Inf),
                          labels = c("18-25","26-35","36-45",
                                     "46-55","56-65","65+")),

    vehicle_age_band = cut(vehicle_age,
                           breaks = c(-1,5,10,15,20,Inf),
                           labels = c("0-5","6-10","11-15",
                                      "16-20","20+"))
  )
```

The derived variables and their purpose are summarised below.

```{r}
#| label: derived-table
#| echo: false
tibble::tribble(
  ~`New Variable`,    ~Formula,                           ~Purpose,
  "loss_ratio",       "total_incurred / total_premium",   "Core profitability metric. Below 1 = profitable, above 1 = loss-making.",
  "claim_frequency",  "total_claims / total_exposure",    "Claims per policy-year. Dividing by exposure makes different policy durations comparable.",
  "severity",         "total_incurred / total_claims",    "Average cost per claim. Only meaningful where total_claims > 0.",
  "has_claim",        "total_claims > 0",                 "Binary flag: did this policy make any claim? Used in association tests.",
  "loss_ratio_bins",  "cut(loss_ratio, ...)",             "Loss ratio split into four categories for ggbarstats association test.",
  "driver_age_band",  "cut(driver_age, ...)",             "Driver age grouped into six brackets for grouped visualisations.",
  "vehicle_age_band", "cut(vehicle_age, ...)",            "Vehicle age grouped into five brackets for grouped visualisations."
) %>%
  knitr::kable(caption = "Derived variables and their analytical purpose")
```

------------------------------------------------------------------------

# 2. Univariate Analysis

## 2.1 Portfolio Composition — Categorical Variables

`geom_bar()` with `geom_text(stat = "count")` displays both count and percentage directly on each bar. `coord_flip()` improves label readability. `patchwork` combines four charts using `|` and `/` operators.

::: panel-tabset
## Portfolio Composition

```{r}
#| label: portfolio-profile
#| fig-height: 6
#| code-fold: true

bar_label <- function(df, var, title) {
  ggplot(df, aes(x = fct_rev({{ var }}))) +
    geom_bar(fill = "#6BAED6") +
    geom_text(
      stat = "count",
      aes(label = paste0(after_stat(count), ",  ",
                         round(after_stat(count) /
                               sum(after_stat(count)) * 100, 1), "%")),
      hjust = -0.1, size = 2.8
    ) +
    coord_flip() +
    scale_y_continuous(expand = expansion(mult = c(0, 0.3))) +
    labs(title = title, x = NULL, y = NULL) +
    theme(axis.text.x  = element_blank(),
          axis.ticks.x = element_blank())
}

p1 <- bar_label(policy, policy_type,       "Policy Type")
p2 <- bar_label(policy, payment_frequency,  "Payment Frequency")
p3 <- bar_label(policy, municipality_type,  "Municipality Type")
p4 <- bar_label(policy %>% filter(!is.na(fuel_type)), fuel_type, "Fuel Type")

(p1 | p2) / (p3 | p4) +
  plot_annotation(
    title      = "Composition of the Motor Insurance Portfolio",
    subtitle   = "Count and percentage of policies across four key categorical variables",
    tag_levels = "A",
    theme = theme(plot.background = element_rect(fill = "#f5f5f5",
                                                 colour = "#f5f5f5"))
  )
```

**Observations**

CC policy type dominates at 57.8% of all policies. Annual payment accounts for over 80%. Inland is the largest geographic segment and Diesel is the dominant fuel type. TP (5,757 policies) and Islands (24,140 policies) are relatively small segments — statistical results for these groups should be treated cautiously.

## Bonus Score

```{r}
#| label: bonus-bar
#| fig-height: 3
#| code-fold: true

ggplot(policy, aes(x = fct_rev(bonus_score))) +
  geom_bar(fill = "#6BAED6") +
  geom_text(stat = "count",
            aes(label = paste0(after_stat(count), ",  ",
                               round(after_stat(count) /
                                     sum(after_stat(count)) * 100, 1), "%")),
            hjust = -0.1, size = 2.5) +
  coord_flip() +
  scale_y_continuous(expand = expansion(mult = c(0, 0.25))) +
  labs(title = "Bonus Score Distribution",
       subtitle = "Good = clean history, Neutral = average, Bad = frequent claimant",
       x = NULL, y = NULL) +
  theme(axis.text.x = element_blank(), axis.ticks.x = element_blank())
```

**Observations**

96.4% of policies are rated Good. Neutral and Bad together account for only 245 policies. Bonus score has limited segmentation power and statistical tests comparing Good vs Bad should be interpreted cautiously due to the extreme imbalance.
:::

------------------------------------------------------------------------

## 2.2 Key Numerical Variables — Distribution

A histogram stacked beneath a boxplot is used. The boxplot reveals outliers and skewness while the histogram shows the full shape. Mean (pink) and median (purple) are annotated. A reusable `hist_box()` function avoids code repetition.

```{r}
#| label: hist-box-fn

hist_box <- function(df, var, var_label) {
  v     <- df %>% filter(!is.na({{ var }})) %>% pull({{ var }})
  m_mn  <- round(mean(v), 1)
  m_med <- round(median(v), 1)

  hist_p <- ggplot(df %>% filter(!is.na({{ var }})),
                   aes(x = {{ var }})) +
    geom_histogram(bins = 25, fill = "#6BAED6", colour = "#f5f5f5") +
    geom_vline(xintercept = m_med, colour = "#595DE5",
               linewidth = 0.7, linetype = "dashed") +
    geom_vline(xintercept = m_mn, colour = "#e0218a",
               linewidth = 0.7, linetype = "dashed") +
    annotate("text", x = m_med, y = Inf,
             label = paste0("Median = ", m_med),
             colour = "#595DE5", hjust = -0.1, vjust = 1.5, size = 3) +
    annotate("text", x = m_mn, y = Inf,
             label = paste0("Mean = ", m_mn),
             colour = "#e0218a", hjust = -0.1, vjust = 3.2, size = 3) +
    labs(x = var_label, y = NULL) +
    theme(axis.text.y = element_blank())

  box_p <- ggplot(df %>% filter(!is.na({{ var }})),
                  aes(y = {{ var }}, x = "")) +
    geom_boxplot(fill = "#6BAED6",
                 outlier.colour = "#d18a7d",
                 outlier.shape  = 16,
                 outlier.size   = 0.8) +
    stat_summary(geom = "point", fun = "mean",
                 colour = "#e0218a", size = 3) +
    stat_boxplot(geom = "errorbar", width = 0.5) +
    coord_flip() +
    theme(axis.text  = element_blank(),
          axis.ticks = element_blank())

  (box_p / hist_p) + plot_layout(heights = c(1, 3.5))
}
```

::: panel-tabset
## Driver Age

```{r}
#| label: dist-driver-age
#| code-fold: true
hist_box(policy, driver_age, "Driver Age (years)")
```

**Observations**

The portfolio is dominated by middle-aged drivers between 40 and 60 years old with a median age of 48. Young drivers under 25 are very few — the insurer appears to either price them out or actively avoid this high-risk group. A small number of elderly drivers in their 80s appear as outliers on the right but are too few to affect the overall picture.

## Vehicle Age

```{r}
#| label: dist-vehicle-age
#| code-fold: true
hist_box(policy, vehicle_age, "Vehicle Age (years)")
```

**Observations**

The typical vehicle is 25 years old — remarkably old by European standards. This is not a new-car portfolio. The distribution shows two clusters, suggesting two types of customers: those with moderately old cars (15 to 20 years) and those keeping very old vehicles on the road (30 to 35 years). Vehicles older than 40 years appear as outliers. Older vehicles tend to have more mechanical failures, creating unpredictable claims costs.

## Total Premium

```{r}
#| label: dist-premium
#| code-fold: true
hist_box(policy, total_premium, "Total Premium (EUR)")
```

**Observations**

Most policies pay between EUR 100 and EUR 600 in annual premium with a median of EUR 258. The histogram shows two visible clusters — a large group of low-premium policies and a smaller group of higher-premium policies. The long right tail with outliers extending beyond EUR 3,000 is visible in the boxplot, while the mean (EUR 297) sits only slightly above the median confirming extreme values are rare.

## Vehicle Value

```{r}
#| label: dist-vehicle-value
#| code-fold: true
hist_box(policy %>% filter(!is.na(vehicle_value)),
         vehicle_value, "Vehicle Value (EUR)")
```

**Observations**

Most vehicles are valued around EUR 25,000 but a small number of luxury or classic cars are worth up to EUR 400,000 — visible as isolated dots far to the right of the boxplot. Despite these extreme outliers, the mean and median are close together confirming the extreme values are very rare. These high-value vehicles are the biggest individual risk in the portfolio — a single claim could generate a loss ratio many times larger than the premium collected.

## Total Incurred

```{r}
#| label: dist-incurred
#| code-fold: true
hist_box(policy %>% filter(total_incurred > 0),
         total_incurred, "Total Incurred (EUR) — Policies With Claims Only")
```

**Observations**

Total incurred is extremely right-skewed even after excluding zero-claim policies. The median incurred amount is modest but the boxplot reveals extreme outliers representing very large individual claims. These extreme claims are the primary driver of portfolio volatility.

## Total Claims

```{r}
#| label: dist-claims
#| code-fold: true
hist_box(policy %>% filter(total_claims > 0),
         total_claims, "Total Claims — Policies With At Least One Claim")
```

**Observations**

Among policies that made at least one claim, the vast majority made exactly one claim. Two or more claims per policy are rare. This confirms that most loss-making events are single-incident rather than repeat claiming behaviour.

## Claim Frequency

```{r}
#| label: dist-claimfreq
#| code-fold: true
hist_box(policy %>% filter(is.finite(claim_frequency),
                           claim_frequency > 0,
                           claim_frequency < 3),
         claim_frequency, "Claim Frequency (claims per year) — Claimants Only")
```

**Observations**

Among policies with at least one claim, the median claim frequency is exactly 1.0 — the typical claimant makes precisely one claim per policy-year. The mean of 1.4 is pulled above the median by a small number of policies making multiple claims, visible as the secondary spike at exactly 2.0. The two distinct spikes at 1.0 and 2.0 reflect the integer nature of claim counts.
:::

------------------------------------------------------------------------

## 2.3 Vehicle Age by Fuel Type — Grouped Density

`geom_density()` overlays two smoothed distributions on the same axis, enabling direct shape comparison between Diesel and Gasoline vehicles.

```{r}
#| label: density-vage-fuel
#| fig-height: 4

ggplot(policy %>% filter(!is.na(fuel_type)),
       aes(x = vehicle_age, colour = fuel_type, fill = fuel_type)) +
  geom_density(alpha = 0.3, linewidth = 0.8) +
  scale_colour_manual(values = c("Diesel" = "#2171B5", "Gasoline" = "#e0218a")) +
  scale_fill_manual(values  = c("Diesel" = "#2171B5", "Gasoline" = "#e0218a")) +
  labs(title = "Gasoline Vehicles Show a Bimodal Age Distribution",
       subtitle = "Kernel density estimate overlaying two groups on the same axis",
       x = "Vehicle Age (years)", y = "Density",
       colour = "Fuel Type", fill = "Fuel Type")
```

**Observations**

Diesel vehicles have a sharp peak at around 18 years, meaning Diesel owners are concentrated in the 10 to 25 year vehicle age range. Gasoline vehicles show a clear bimodal distribution with two peaks with one at around 25 years and a second at around 35 years, suggesting two distinct customer segments. The two curves overlap closely between 20 and 30 years but diverge significantly beyond 30 where Gasoline has considerably more mass, meaning Gasoline vehicles are more likely to be very old and may generate higher maintenance-related claims.

------------------------------------------------------------------------

## 2.4 Vehicle Value by Policy Type — Violin Plot

A violin plot shows the full density shape per group. An inner boxplot is overlaid with `geom_boxplot(width = 0.1)` to retain median and IQR information. Mean is added via `stat_summary()`.

```{r}
#| label: violin-value-ptype
#| fig-height: 4

ggplot(policy %>% filter(!is.na(vehicle_value)),
       aes(x = policy_type, y = vehicle_value)) +
  geom_violin(fill = "#6BAED6", alpha = 0.7) +
  geom_boxplot(width = 0.1, fill = "white", outlier.shape = NA) +
  stat_summary(geom = "point", fun = "mean",
               colour = "#e0218a", size = 3) +
  labs(title = "Comprehensive Policies Cover Higher-Value Vehicles",
       subtitle = "Violin plot with inner boxplot and mean point (pink)",
       x = "Policy Type", y = "Vehicle Value (EUR)")
```

**Observations**

All five policy types have most vehicle values clustered near zero, reflected in the flat and wide violin bases. COMP_E and COMP_N stand out with higher mean points and slightly less flat bases, indicating their vehicles are more spread across a wider value range. CC has the longest upper tail reaching EUR 350,000 due to its large policy count rather than genuinely higher vehicle values. The overall pattern confirms that comprehensive policyholders tend to own more expensive vehicles, which directly drives higher claim costs when damage occurs.

------------------------------------------------------------------------

## 2.5 Loss Ratio Distribution — The Headline Metric

Because 88% of policies have zero incurred losses, the chart is restricted to policies where both `total_claims > 0` and `total_incurred > 0`. The excluded proportion is calculated dynamically and annotated on the chart.

```{r}
#| label: loss-ratio-overall
#| fig-height: 4

pct_excluded <- scales::percent(
  1 - mean(policy$total_claims > 0 & policy$total_incurred > 0, na.rm = TRUE),
  accuracy = 0.1)

claims_only <- policy %>%
  filter(total_claims > 0, total_incurred > 0,
         is.finite(loss_ratio), loss_ratio < 5)

max_y <- claims_only %>%
  pull(loss_ratio) %>%
  cut(breaks = seq(0, 5, length.out = 41)) %>%
  table() %>% max()

ggplot(claims_only, aes(x = loss_ratio)) +
  geom_histogram(bins = 40, fill = "#6BAED6", colour = "#f5f5f5") +
  geom_vline(xintercept = 1, colour = "firebrick",
             linetype = "dashed", linewidth = 0.8) +
  annotate("text", x = 1.08, y = max_y * 0.6,
           label = "Break-even\n(LR = 1)",
           colour = "firebrick", hjust = 0, size = 3.5) +
  annotate("text", x = 3, y = max_y * 0.9,
           label = paste0(pct_excluded, " of policies\nexcluded\n(no paid claims)"),
           colour = "#2171B5", hjust = 0, size = 3.5) +
  labs(title = "Among Policies With Paid Claims, Many Exceed Break-Even",
       subtitle = paste0("Showing ", scales::comma(nrow(claims_only)),
                         " policies with at least one paid claim (",
                         pct_excluded, " excluded)"),
       x = "Loss Ratio  (incurred / premium)",
       y = "Number of Policies With a Claim")
```

**Observations**

Of the 32,637 policies with at least one paid claim, more sit below break-even than above it — meaning the majority of claimants are still profitable for the insurer. However the distribution does not drop sharply after LR = 1. Instead it declines gradually and remains substantial all the way to LR = 5, meaning loss-making claims are not rare outliers but a persistent and sizable tail. The portfolio profits from the 88.1% who never claim, but remains exposed to a long tail of severe claimants that is difficult to price away.

------------------------------------------------------------------------

# 3. Bivariate Analysis — What Drives Profitability and Volatility?

For each hypothesis, a visual is paired with a statistical test. Non-parametric tests (`type = "np"`) are used throughout as the loss ratio is heavily right-skewed and violates normality assumptions.

------------------------------------------------------------------------

## 3.1 Policy Type — The Primary Driver of Loss Ratio

Policy type is examined first as it is the most actionable pricing variable. Three complementary charts are used: a boxplot for the distribution shape, a statistical test for significance, and a claim rate chart to decompose the mechanism.

::: panel-tabset
## Boxplot — Loss Ratio by Policy Type

```{r}
#| label: notched-box-ptype
#| fig-height: 4
#| code-fold: true

policy %>%
  filter(is.finite(loss_ratio), loss_ratio < 5) %>%
  ggplot(aes(x = policy_type, y = loss_ratio)) +
  geom_boxplot(fill = "#6BAED6",
               outlier.shape = NA) +
  stat_summary(geom = "point", fun = "mean",
               colour = "#e0218a", size = 3) +
  geom_hline(yintercept = 1, colour = "firebrick",
             linetype = "dashed", linewidth = 0.7) +
  annotate("text", x = 1, y = 1.05,
           label = "Break-even", colour = "firebrick", size = 3) +
  coord_cartesian(ylim = c(0, 2)) +
  labs(title = "COMP_N Has the Highest Mean Loss Ratio Across All Policy Types",
       subtitle = "Boxplot zoomed to 0-2. Pink point = mean. Dashed = break-even.",
       x = "Policy Type", y = "Loss Ratio (zoomed to 0-2)")
```

**Observations**

All five policy types have medians near zero and means well below break-even, meaning every segment is profitable on average. A clear gradient emerges as TP and TPG means sit closest to zero while COMP_N's mean at approximately 0.5 is the highest of all types. COMP_N is also the only segment with a visible IQR box, indicating meaningfully more spread in its loss ratios. While COMP_N remains profitable on average, its wider distribution and higher mean signal that it carries the most risk.

## Statistical Test — Kruskal-Wallis

```{r}
#| label: lr-by-ptype
#| fig-height: 5
#| code-fold: true

ggbetweenstats(
  data     = policy %>% filter(is.finite(loss_ratio), loss_ratio < 5),
  x        = policy_type,
  y        = loss_ratio,
  type     = "np",
  pairwise.comparisons = TRUE,
  pairwise.display     = "s",
  p.adjust.method      = "fdr",
  messages  = FALSE,
  title     = "Kruskal-Wallis Test — Loss Ratio Differs Significantly Across Policy Types",
  xlab      = "Policy Type",
  ylab      = "Loss Ratio"
)
```

**Observations**

All five policy types have a median loss ratio of exactly 0.00, meaning more than half of policies in every segment make no claims at all. COMP_N is the only type with a visible interquartile range box, confirming it has the most spread. The Kruskal-Wallis test confirms overall differences are statistically significant (p = 2.23e-308), and all pairwise FDR-corrected comparisons show significant p-values. Hence we can say that every policy type is genuinely different from every other. Despite identical medians, COMP_N carries far more variability in loss outcomes.

## Claim Rate by Policy Type

```{r}
#| label: claim-rate-ptype
#| fig-height: 4
#| code-fold: true

policy %>%
  group_by(policy_type) %>%
  summarise(claim_rate = mean(has_claim) * 100, n = n()) %>%
  ggplot(aes(x = policy_type, y = claim_rate)) +
  geom_col(fill = "#6BAED6", width = 0.6) +
  geom_text(aes(label = paste0(round(claim_rate, 1), "%",
                               "\n(n=", scales::comma(n), ")")),
            vjust = -0.3, size = 3.5) +
  scale_y_continuous(limits = c(0, 50), labels = NULL) +
  labs(title = "COMP_N Policies Claim Most Frequently — Over 3x the TP Rate",
       subtitle = "Percentage of policies with at least one claim by policy type",
       x = "Policy Type", y = NULL) +
  theme(axis.ticks.y = element_blank())
```

**Observations**

Claim rate rises steadily as cover becomes more comprehensive. TP has the lowest claim rate at 6.9% meaning that customers with basic third-party only cover tend to absorb small damages themselves rather than claim. COMP_N has the highest at 31.3%. The jump from COMP_E (18.5%) to COMP_N (31.3%) is particularly striking given that these two products offer the same breadth of cover. The only difference is the excess. This single design feature nearly doubles the claim rate, directly explaining why COMP_N generates higher loss ratios than COMP_E.
:::

------------------------------------------------------------------------

## 3.2 Loss Distribution by Policy Type and Bonus Score — Ridgeline Plots

`stat_density_ridges()` with ECDF gradient and quantile lines reveals how loss ratio distributions differ in shape across segments, beyond what boxplots can show.

::: panel-tabset
## By Policy Type (ECDF fill)

```{r}
#| label: ridge-ecdf-ptype
#| code-fold: true
#| fig-height: 5

policy %>%
  filter(is.finite(loss_ratio),
         loss_ratio < 5) %>%
  ggplot(aes(x = loss_ratio,
             y = fct_rev(policy_type),
             fill = 0.5 - abs(0.5 - after_stat(ecdf)))) +
  stat_density_ridges(
    geom      = "density_ridges_gradient",
    calc_ecdf = TRUE,
    alpha     = 0.8,
    scale     = 1.5
  ) +
  scale_fill_viridis_c(name = "Probability", option = "C") +
  geom_vline(xintercept = 1, colour = "firebrick",
             linetype = "dashed", linewidth = 0.7) +
  coord_cartesian(xlim = c(0, 1)) +
  labs(title = "TP and TPG Are Most Predictable — COMP_N Most Volatile",
       subtitle = "ECDF gradient: brighter colour = most probable values. All policies included. Dashed = break-even.",
       x = "Loss Ratio", y = NULL) +
  theme_ridges(font_size = 12) +
  theme(plot.background = element_rect(fill = "#f5f5f5", colour = "#f5f5f5"))
```

**Observations**

All five policy types show a bright yellow peak concentrated near zero, confirming that the most probable outcome for any policy is zero or near-zero loss ratio. TP and TPG have the tallest and most concentrated peaks, meaning their loss ratios are highly predictable. CC and COMP_E peaks are slightly lower and wider. COMP_N has the flattest and widest peak of all five. Its bright yellow region is less intense and the distribution spreads further toward break-even, confirming that loss ratio outcomes are the most variable and hardest to predict as cover becomes more comprehensive.

## By Bonus Score (Quantile lines)

```{r}
#| label: ridge-quantile-bonus
#| code-fold: true
#| fig-height: 4

policy %>%
  filter(is.finite(loss_ratio),
         loss_ratio < 5) %>%
  ggplot(aes(x = loss_ratio,
             y = fct_rev(bonus_score),
             fill = factor(after_stat(quantile)))) +
  stat_density_ridges(
    geom           = "density_ridges_gradient",
    calc_ecdf      = TRUE,
    quantiles      = c(0.5, 0.75, 0.90),
    quantile_lines = TRUE
  ) +
  scale_fill_viridis_d(name = "Quantiles", alpha = 0.8) +
  geom_vline(xintercept = 1, colour = "firebrick",
             linetype = "dashed", linewidth = 0.7) +
  coord_cartesian(xlim = c(0, 2)) +
  labs(title = "Bad Bonus Score Pushes the Loss Tail Beyond Break-Even",
       subtitle = "All policies included. Lines show 50th, 75th and 90th percentiles. Dashed = break-even.",
       x = "Loss Ratio", y = NULL) +
  theme_ridges(font_size = 12) +
  theme(plot.background = element_rect(fill = "#f5f5f5", colour = "#f5f5f5"))
```

**Observations**

All three bonus score groups show a dominant peak near zero. For Good customers, the 90th percentile line falls well below break-even at around 0.3, meaning 90% of Good policies have a loss ratio below 0.3. For Bad customers, the 90th percentile line crosses break-even at LR = 1, meaning the top 10% of Bad score policies are loss-making. The yellow region (top 10%) for Bad extends all the way to LR = 2, confirming that bonus score differentiates tail risk rather than typical outcomes.
:::

------------------------------------------------------------------------

## 3.3 Bonus Score — Statistical Evidence

`ggbetweenstats()` and `ggbarstats()` formally test whether bonus score is statistically associated with loss ratio outcomes.

::: panel-tabset
## Kruskal-Wallis Test

```{r}
#| label: lr-by-bonus
#| fig-height: 5
#| code-fold: true

ggbetweenstats(
  data     = policy %>% filter(is.finite(loss_ratio), loss_ratio < 5),
  x        = bonus_score,
  y        = loss_ratio,
  type     = "np",
  messages  = FALSE,
  title     = "Bad Claims History Is Associated With Higher Loss Ratio Variability",
  xlab      = "Bonus Score",
  ylab      = "Loss Ratio"
)
```

**Observations**

All three bonus score groups have a median loss ratio of exactly 0.00. However the violin shapes reveal a clear difference where the Good group has the narrowest violin meaning most policies cluster tightly near zero with little variation. Neutral and Bad groups have progressively wider violins showing that loss ratio outcomes become more spread out as claims history worsens. The Kruskal-Wallis test confirms these distributional differences are statistically significant (p = 6.99e-69), with all three pairwise comparisons also significant. Bonus score is a meaningful signal of risk and should be used as a pricing variable.

## Association Test — ggbarstats

```{r}
#| label: ggbar-bonus-lrbins
#| code-fold: true
#| fig-height: 5

ggbarstats(
  data  = policy %>% filter(!is.na(loss_ratio_bins), !is.na(bonus_score)),
  x     = loss_ratio_bins,
  y     = bonus_score,
  title = "Is Bonus Score Associated With Loss-Making Outcomes?"
)
```

**Observations**

Across all three bonus score groups, the majority of policies make no claims at all with 88% for Good, 85% for Neutral and 81% for Bad. However the loss-making proportion (Above 1) grows steadily from 9% for Good to 13% for Bad, meaning Bad customers are 44% more likely to generate a loss-making policy. The chi-squared test confirms this association is statistically significant (p = 1.35e-94), though the small Cramer's V of 0.03 indicates the effect size is modest. Bonus score shifts the probabilities but does not dramatically change typical outcomes.
:::

------------------------------------------------------------------------

## 3.4 Confidence Intervals — Uncertainty of Group Means

The standard error is computed as `se = sd / sqrt(n-1)`. The 99% confidence interval is `mean ± 2.58 × se`, making the calculation fully transparent.

```{r}
#| label: manual-errorbars
#| code-fold: true
#| fig-height: 4

my_sum <- policy %>%
  filter(is.finite(loss_ratio), loss_ratio < 5) %>%
  group_by(policy_type) %>%
  summarise(n    = n(),
            mean = mean(loss_ratio, na.rm = TRUE),
            sd   = sd(loss_ratio,   na.rm = TRUE)) %>%
  mutate(se = sd / sqrt(n - 1))

ggplot(my_sum) +
  geom_errorbar(aes(x = policy_type,
                    ymin = mean - 2.58 * se,
                    ymax = mean + 2.58 * se),
                width = 0.2, colour = "grey40") +
  geom_point(aes(x = policy_type, y = mean),
             colour = "#e0218a", size = 4) +
  geom_hline(yintercept = 1, colour = "firebrick",
             linetype = "dashed", linewidth = 0.7) +
  annotate("text", x = 0.6, y = 1.05,
           label = "Break-even", colour = "firebrick", size = 3.2) +
  labs(title = "99% Confidence Intervals for Policy Type",
       subtitle = "Mean loss ratio +/- 2.58 SE (99% CI). Pink point = mean.",
       x = "Policy Type", y = "Mean Loss Ratio")
```

**Observations**

All five policy types have mean loss ratios well below the break-even line of 1.0, confirming the entire portfolio is profitable on average. The confidence intervals for TPG, CC and COOMP_E are extremely tight, meaning their profitability is highly certain but COMP_E sits higher at around 0.25 . COMP_N stands clearly apart with a mean of approximately 0.47 and the widest confidence interval of all five, showing that not only does it generate the highest average loss ratio, it is also the most uncertainty around that estimate, making it the hardest segment to price precisely.

------------------------------------------------------------------------

## 3.5 Loss Ratio Trend Across Years

`geom_line()` tracks group means across years. `geom_text()` labels each point directly to avoid a separate legend lookup.

```{r}
#| label: lr-by-year
#| fig-height: 4

policy %>%
  filter(is.finite(loss_ratio), loss_ratio < 5) %>%
  group_by(year, policy_type) %>%
  summarise(mean_lr = mean(loss_ratio, na.rm = TRUE), .groups = "drop") %>%
  ggplot(aes(x = year, y = mean_lr,
             colour = policy_type, group = policy_type)) +
  geom_line(linewidth = 1) +
  geom_point(size = 3) +
  geom_text(aes(label = round(mean_lr, 2)),
            vjust = -0.8, size = 2.8) +
  geom_hline(yintercept = 1, colour = "firebrick",
             linetype = "dashed", linewidth = 0.7) +
  scale_x_continuous(breaks = unique(policy$year)) +
  scale_colour_brewer(palette = "Set1") +
  labs(title = "COMP_N Is Deteriorating Rapidly — Heading Toward Break-Even",
       subtitle = "Mean loss ratio per year by policy type. Dashed = break-even.",
       x = "Year", y = "Mean Loss Ratio", colour = "Policy Type")
```

**Observations**

All five types remain well below break-even throughout all three years, confirming the portfolio is consistently profitable. TP, TPG and CC remain flat and stable, indicating these segments are reliably priced. COMP_E shows a gradual upward trend from 0.19 to 0.26. COMP_N is the most concerning as it rose sharply from 0.30 in 2022 to 0.40 in 2023 and further to 0.61 in 2024. This is not a one-year anomaly but a consistent three-year deterioration that signals an urgent need for premium repricing in the COMP_N segment.

------------------------------------------------------------------------

## 3.6 Driver Age and Claim Frequency

Claim frequency is examined by driver age group using a jitter plot for the raw distribution and a Kruskal-Wallis test for statistical confirmation.

::: panel-tabset
## Jitter Plot — Claim Frequency by Age Group

```{r}
#| label: jitter-freq-ageband
#| fig-height: 4.5
#| code-fold: true

overall_mean_freq <- mean(policy$claim_frequency[is.finite(policy$claim_frequency)],
                          na.rm = TRUE)

cap <- quantile(policy$claim_frequency[is.finite(policy$claim_frequency)],
                0.95, na.rm = TRUE)

policy %>%
  filter(!is.na(driver_age_band),
         is.finite(claim_frequency),
         claim_frequency < cap) %>%
  ggplot(aes(x = claim_frequency,
             y = driver_age_band,
             colour = driver_age_band)) +
  geom_jitter(size = 0.5, alpha = 0.08,
              height = 0.3,
              width = 0.05) +
  stat_summary(fun = "median", geom = "point",
               colour = "#595DE5", size = 4, alpha = 0.9) +
  geom_vline(xintercept = overall_mean_freq,
             colour = "firebrick", linetype = "dashed", linewidth = 0.7) +
  annotate("text", x = overall_mean_freq + 0.02, y = 6.4,
           label = paste0("Portfolio mean = ",
                          round(overall_mean_freq, 3)),
           colour = "firebrick", size = 3.2, hjust = 0) +
  scale_colour_brewer(palette = "Blues") +
  labs(title = "Claim Frequency by Each Age Group",
       subtitle = "Each dot = one policy (jittered). Blue point = median. Dashed = portfolio mean.",
       x = "Claim Frequency (claims per year)",
       y = "Driver Age Group") +
  theme(legend.position = "none")
```

**Observations**

The chart shows two distinct clusters for every age group with a large block at exactly 0 (no claims) and a second cluster starting at exactly 1 (at least one claim per year). All age group medians sit at 0, confirming the majority of policies in every age group make no claims. Among those that do claim, older groups (46 and above) show slightly denser dot bands at claim frequency = 1. The 18-25 group has almost no dots due to its very small sample size of only 1,616 policies, making any comparison unreliable.

## Statistical Test — Kruskal-Wallis

```{r}
#| label: ggbetween-claimfreq-age
#| code-fold: true
#| fig-height: 5

ggbetweenstats(
  data     = policy %>%
               filter(!is.na(driver_age_band),
                      is.finite(claim_frequency),
                      claim_frequency < 2),
  x        = driver_age_band,
  y        = claim_frequency,
  type     = "np",
  pairwise.comparisons = TRUE,
  pairwise.display     = "s",
  p.adjust.method      = "fdr",
  centrality.label.args = list(size = 2,
                                nudge_x = 0.4,
                                nudge_y = 0.1),
  messages  = FALSE,
  title     = "Does Claim Frequency Differ Across Driver Age Groups?",
  xlab      = "Driver Age Group",
  ylab      = "Claim Frequency (claims per year)"
)
```

**Observations**

The Kruskal-Wallis test confirms that claim frequency differences across driver age groups are statistically significant (p = 3.67e-43). All medians are 0.00 — the difference lies in the distribution tails. Most pairwise comparisons are significant, confirming that age genuinely affects claim behaviour. Age-based pricing adjustments are actuarially justified by this finding.

## Correlation — Driver Age vs Claim Frequency

```{r}
#| label: ggscatter-age-freq
#| code-fold: true
#| fig-height: 5

ggscatterstats(
  data     = policy %>% filter(is.finite(claim_frequency),
                               claim_frequency < 3),
  x        = driver_age,
  y        = claim_frequency,
  marginal = FALSE,
  title    = "Older Drivers Claim Slightly Less Often",
  xlab     = "Driver Age (years)",
  ylab     = "Claim Frequency (claims per year)"
)
```

**Observations**

The Spearman correlation is negative and statistically significant, showing that older drivers claim less. The effect size is small, but with 354,140 rows even tiny real effects are detectable. This further supports age-based pricing differentiation.
:::

------------------------------------------------------------------------

## 3.7 Loss Ratio by Fuel Type — Raincloud Plot

The raincloud design (`stat_halfeye()` + `geom_boxplot()` + `stat_dots()`) shows three complementary views simultaneously: density shape, five-number summary, and individual policy dots. Analysis is restricted to policies with paid claims to avoid the zero-spike problem.

::: panel-tabset
## Diesel

```{r}
#| label: raincloud-diesel
#| code-fold: true
#| fig-height: 4

ggplot(policy %>% filter(fuel_type == "Diesel",
                         total_claims > 0,
                         total_incurred > 0,
                         is.finite(loss_ratio),
                         loss_ratio < 3),
       aes(x = fuel_type, y = loss_ratio)) +
  stat_halfeye(adjust = 0.5, justification = -0.2,
               .width = 0, point_colour = NA,
               fill = "#2171B5", alpha = 0.7) +
  geom_boxplot(width = 0.15, outlier.shape = NA,
               alpha = 0.6, fill = "#2171B5") +
  stat_dots(side = "left", justification = 1.1,
            binwidth = NA,
            alpha = 0.3) +
  coord_flip() +
  labs(title = "Diesel — Loss Ratio Distribution",
       subtitle = "Policies with paid claims only. Density (top) + boxplot (middle) + dots (bottom). Capped at LR = 3.",
       x = NULL, y = "Loss Ratio") +
  theme_economist()
```

## Gasoline

```{r}
#| label: raincloud-gasoline
#| code-fold: true
#| fig-height: 4

ggplot(policy %>% filter(fuel_type == "Gasoline",
                         total_claims > 0,
                         total_incurred > 0,
                         is.finite(loss_ratio),
                         loss_ratio < 3),
       aes(x = fuel_type, y = loss_ratio)) +
  stat_halfeye(adjust = 0.5, justification = -0.2,
               .width = 0, point_colour = NA,
               fill = "#e0218a", alpha = 0.7) +
  geom_boxplot(width = 0.15, outlier.shape = NA,
               alpha = 0.6, fill = "#e0218a") +
  stat_dots(side = "left", justification = 1.1,
            binwidth = NA,
            alpha = 0.3) +
  coord_flip() +
  labs(title = "Gasoline — Loss Ratio Distribution",
       subtitle = "Policies with paid claims only. Density (top) + boxplot (middle) + dots (bottom). Capped at LR = 3.",
       x = NULL, y = "Loss Ratio") +
  theme_economist()
```
:::

**Observations**

Both Diesel and Gasoline show nearly identical flat distributions across the entire loss ratio range from 0 to 3, confirming that once a claim occurs, its cost relative to premium is highly unpredictable for both fuel types. Both medians sit close to 1, meaning the typical claimant costs the insurer roughly what it collected in premium. The boxplot for Gasoline is slightly wider than Diesel, suggesting marginally more variability in Gasoline claim outcomes. Fuel type does not meaningfully change how costly claims are as the difference between the two groups is more likely driven by claim frequency than claim severity.

------------------------------------------------------------------------

## 3.8 Vehicle Age vs Loss Ratio — Scatter With Regression

`geom_smooth(method = lm)` adds a linear regression line with a confidence band. `geom_point()` with `alpha = 0.15` handles overplotting.

```{r}
#| label: scatter-vage-lr
#| fig-height: 4.5

policy %>%
  filter(is.finite(loss_ratio), loss_ratio < 5) %>%
  slice_sample(n = 5000) %>%
  ggplot(aes(x = vehicle_age, y = loss_ratio)) +
  geom_point(alpha = 0.15, size = 0.8, colour = "#6BAED6") +
  geom_smooth(method = lm, linewidth = 1,
              colour = "#e0218a", se = TRUE) +
  geom_hline(yintercept = 1, colour = "firebrick",
             linetype = "dashed", linewidth = 0.7) +
  annotate("text", x = 55, y = 1.1,
           label = "Break-even", colour = "firebrick", size = 3.2) +
  labs(title = "No Strong Linear Relationship Between Vehicle Age and Loss Ratio",
       subtitle = "Scatter with linear regression line (pink). Random sample of 5,000 policies.",
       x = "Vehicle Age (years)", y = "Loss Ratio")
```

**Observations**

The regression line is nearly flat, confirming that vehicle age alone does not drive loss ratio. The wide confidence band reflects the enormous individual variation in loss outcomes. Policy type and bonus score are stronger and more actionable drivers than vehicle age.

------------------------------------------------------------------------

## 3.9 Geography — Does Municipality Type Matter?

`facet_wrap()` produces small multiples with an identical x-axis scale, allowing shape comparison across geography without scale distortion.

```{r}
#| label: facet-lr-municipality
#| fig-height: 4

policy %>%
  filter(total_claims > 0, total_incurred > 0,
         is.finite(loss_ratio), loss_ratio < 5) %>%
  ggplot(aes(x = loss_ratio)) +
  geom_histogram(bins = 30, fill = "#6BAED6", colour = "#f5f5f5") +
  geom_vline(xintercept = 1, colour = "firebrick",
             linetype = "dashed", linewidth = 0.7) +
  facet_wrap(~ municipality_type) +
  labs(title = "Loss Ratio Shape Is Consistent Across Municipality Types",
       subtitle = "Policies with paid claims only. Dashed line = break-even (LR = 1).",
       x = "Loss Ratio", y = "Number of Policies")
```

**Observations**

All three municipality panels share the same right-skewed loss ratio shape. The Islands panel is smaller due to fewer policies but shows no unique distributional characteristics. The proportion of bars above versus below break-even looks similar across all three panels, confirming that geography is not a primary driver of loss ratio differences and is not a useful pricing lever in this portfolio.

------------------------------------------------------------------------

# 4. Bonus — Interactive Views

## 4.1 Interactive Scatter — Claim Frequency vs Loss Ratio

`geom_point_interactive()` with `tooltip` and `data_id` enables hover detail and group highlighting. `opts_hover_inv()` dims all other points when a group is hovered.

```{r}
#| label: interactive-scatter

policy_plot <- policy %>%
  filter(is.finite(loss_ratio), is.finite(claim_frequency),
         loss_ratio < 5, claim_frequency < 3) %>%
  slice_sample(n = 5000) %>%
  mutate(tooltip = paste0(
    "Policy Type: ", policy_type,
    "\nFuel: ", fuel_type,
    "\nDriver Age: ", driver_age,
    "\nLoss Ratio: ", round(loss_ratio, 2),
    "\nClaim Freq: ", round(claim_frequency, 2)
  ))

p <- ggplot(policy_plot,
            aes(x = claim_frequency, y = loss_ratio,
                colour = policy_type)) +
  geom_point_interactive(aes(tooltip = tooltip,
                             data_id  = policy_type),
                         alpha = 0.5, size = 1.5) +
  geom_hline(yintercept = 1, colour = "firebrick", linetype = "dashed") +
  scale_colour_brewer(palette = "Set1") +
  labs(title = "Hover to Explore Individual Policies — Click Policy Type to Isolate",
       x = "Claim Frequency (claims per year)",
       y = "Loss Ratio", colour = "Policy Type") +
  theme(plot.background = element_rect(fill = "#f5f5f5", colour = "#f5f5f5"))

girafe(ggobj = p, width_svg = 9, height_svg = 5,
       options = list(
         opts_hover(css     = "stroke:black; stroke-width:1.5;"),
         opts_hover_inv(css = "opacity:0.1;")
       ))
```

**Observations**

COMP_N points cluster most densely above the break-even line, confirming the bivariate findings interactively. Hovering reveals that high-loss-ratio policies tend to have low claim frequency — one severe claim rather than many small ones drives the extreme outcomes.

------------------------------------------------------------------------

## 4.2 Vehicle Brand vs Loss Ratio

`ggplotly()` makes the scatter interactive — hover over any dot to see the brand name, earned premium and loss ratio.

```{r}
#| label: funnel-brand
#| fig-height: 5

brand_sum <- policy %>%
  group_by(vehicle_brand) %>%
  summarise(incurred   = sum(total_incurred, na.rm = TRUE),
            premium    = sum(total_premium,  na.rm = TRUE),
            loss_ratio = incurred / premium,
            .groups    = "drop") %>%
  filter(premium > 0)

p <- ggplot(brand_sum, aes(x = premium, y = loss_ratio,
                            label = vehicle_brand)) +
  geom_point(colour = "#6BAED6", size = 3, alpha = 0.7) +
  geom_hline(yintercept = mean(brand_sum$loss_ratio),
             linetype = "dashed", colour = "black") +
  geom_text(aes(label = vehicle_brand),
            hjust = -0.1, size = 1.6, check_overlap = TRUE) +
  labs(title = "Vehicle Brand vs Loss Ratio — No Brand Stands Out as Problematic",
       x = "Earned Premium (EUR)",
       y = "Loss Ratio") +
  theme_minimal()

ggplotly(p, tooltip = c("label", "x", "y"))
```

**Observations**

The 11 vehicle brands cluster within a narrow loss ratio range of 0.66 to 0.78, which is a difference of only 0.12 across all brands. Opel, Toyota and Seat sit above the portfolio average while Audi, Ford, Renault and Peugeot perform better than average. However the gap is small enough that brand alone is not a meaningful pricing lever. No brand stands out as genuinely problematic.

------------------------------------------------------------------------

## 4.3 Interactive Data Table

`DT::datatable()` with `filter = "top"` adds column-level search boxes. Any reader can verify all findings by filtering the underlying data directly.

```{r}
#| label: dt-table

policy %>%
  filter(is.finite(loss_ratio)) %>%
  select(insured_id, year, policy_type, fuel_type,
         driver_age, vehicle_age, total_premium,
         total_incurred, loss_ratio, claim_frequency, has_claim) %>%
  mutate(across(where(is.numeric), ~ round(., 3))) %>%
  DT::datatable(caption = "Full Cleaned Dataset — Search and Filter",
                filter  = "top",
                class   = "compact",
                options = list(pageLength = 10, scrollX = TRUE))
```

------------------------------------------------------------------------

# 5. Summary of Findings

```{r}
#| label: findings-table
#| echo: false

tibble::tribble(
  ~Finding, ~Evidence, ~Implication,
  "Portfolio is profitable overall",
    "Mean LR well below 1 across all segments; 88% of policies make no claims",
    "The book earns more than it pays but tail claims drive all the risk",
  "COMP_N has the highest loss ratio and is deteriorating",
    "Mean LR = 0.47, rising from 0.30 in 2022 to 0.61 in 2024; claim rate 31.3%",
    "Urgent repricing of COMP_N premiums required",
  "COMP_N is the most volatile segment",
    "Widest IQR box, widest 99% confidence interval, flattest ridgeline",
    "Makes COMP_N the hardest to price precisely — highest uncertainty",
  "Bad bonus score predicts higher tail risk",
    "90th percentile for Bad crosses break-even; chi-squared p = 1.35e-94",
    "Bonus score should be used aggressively in pricing tiers",
  "Driver age is negatively correlated with claim frequency",
    "Kruskal-Wallis p = 3.67e-43; Spearman correlation negative and significant",
    "Age-based pricing adjustments are actuarially justified",
  "Geography and vehicle age are weak drivers",
    "Faceted histogram shows identical shape; regression slope nearly flat",
    "Policy type and bonus score are stronger and more actionable pricing variables"
) %>%
  knitr::kable(
    col.names = c("Finding", "Visual Evidence", "Implication"),
    caption   = "Summary of Visual Analysis Findings"
  )
```
