---
title: "Hands-on_Ex09a"
author: "Huang Zhanyi"
date: "June 18, 2026"
date-modified: "last-modified"
execute: 
  eval: true
  echo: true
  warning: false
  freeze: true
---

# 21 Choropleth Mapping with R

## 21.1 Overview

Choropleth mapping involves the symbolisation of enumeration units, such as countries, provinces, states, counties or census units, using area patterns or graduated colors.

## 21.2 Getting Started

Loading required packages:

```{r}
pacman::p_load(sf, tmap, tidyverse)
```

## 21.3 Importing Data into R

### 21.3.2 Importing Geospatial Data into R

```{r}
mpsz <- st_read(dsn = "data/geospatial", 
                layer = "MP14_SUBZONE_WEB_PL")
```

Examine the content of mpsz:

```{r}
mpsz
```

### 21.3.3 Importing Attribute Data into R

```{r}
popdata <- read_csv("data/aspatial/respopagesextod2011to2020.csv")
```

### 21.3.4 Data Preparation

#### 21.3.4.1 Data wrangling

```{r}
popdata2020 <- popdata %>%
  filter(Time == 2020) %>%
  group_by(PA, SZ, AG) %>%
  summarise(`POP` = sum(`Pop`)) %>%
  ungroup() %>%
  pivot_wider(names_from=AG, 
              values_from=POP) %>%
  mutate(YOUNG = rowSums(.[3:6])
         +rowSums(.[12])) %>%
  mutate(`ECONOMY ACTIVE` = rowSums(.[7:11])+
         rowSums(.[13:15]))%>%
  mutate(`AGED`=rowSums(.[16:21])) %>%
  mutate(`TOTAL`=rowSums(.[3:21])) %>%  
  mutate(`DEPENDENCY` = (`YOUNG` + `AGED`)
         /`ECONOMY ACTIVE`) %>%
  select(`PA`, `SZ`, `YOUNG`, 
         `ECONOMY ACTIVE`, `AGED`, 
         `TOTAL`, `DEPENDENCY`)
```

#### 21.3.4.2 Joining the attribute data and geospatial data

Convert PA and SZ to uppercase:

```{r}
popdata2020 <- popdata2020 %>%
  mutate(across(c(PA, SZ), toupper)) %>%
  filter(`ECONOMY ACTIVE` > 0)
```

Join geographical data and attribute table:

```{r}
mpsz_pop2020 <- left_join(mpsz, popdata2020,
                          by = c("SUBZONE_N" = "SZ"))
```

Save the result:

```{r}
write_rds(mpsz_pop2020, "data/rds/mpszpop2020.rds")
```

## 21.4 Choropleth Mapping Geospatial Data Using tmap

### 21.4.1 Plotting a choropleth map quickly by using qtm()

```{r}
tmap_mode("plot")
qtm(mpsz_pop2020, 
    fill = "DEPENDENCY")
```

### 21.4.2 Creating a choropleth map by using tmap's elements

#### 21.4.2.1 Drawing a base map

```{r}
tm_shape(mpsz_pop2020) +
  tm_polygons()
```

#### 21.4.2.2 Drawing a choropleth map using tm_polygons()

```{r}
tm_shape(mpsz_pop2020)+
  tm_polygons("DEPENDENCY")
```

#### 21.4.2.3 Drawing a choropleth map using tm_fill() and tm_border()

Using tm_fill() alone:

```{r}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY")
```

Adding borders:

```{r}
tm_shape(mpsz_pop2020)+
  tm_polygons(fill = "DEPENDENCY") +
  tm_borders(lwd = 0.01,  
             fill_alpha = 0.1)
```

### 21.4.3 Data classification methods of tmap

#### 21.4.3.1 Plotting choropleth maps with built-in classification methods

Using Jenks classification:

```{r}
tm_shape(mpsz_pop2020)+
  tm_polygons("DEPENDENCY",
              fill.scale = tm_scale_intervals(
                style = "jenks",
                n = 5)) +
  tm_borders(fill_alpha = 0.5)
```

Using equal classification:

```{r}
tm_shape(mpsz_pop2020)+
  tm_polygons("DEPENDENCY",
              fill.scale = tm_scale_intervals(
                style = "equal",
                n = 5)) +
  tm_borders(fill_alpha = 0.5)
```

#### 21.4.3.2 Plotting choropleth map with custom breaks

Get descriptive statistics:

```{r}
summary(mpsz_pop2020$DEPENDENCY)
```

Plot with custom breaks:

```{r}
tm_shape(mpsz_pop2020)+
  tm_polygons("DEPENDENCY",
              breaks = c(0, 0.60, 0.70, 0.80, 0.90, 1.00)) +
  tm_borders(fill_alpha = 0.5)
```

### 21.4.4 Colour Scheme

#### 21.4.4.1 Using ColourBrewer palette

Using brewer.greens:

```{r}
tm_shape(mpsz_pop2020)+
  tm_polygons("DEPENDENCY",
              fill.scale = tm_scale_intervals(
                style = "quantile",
                n = 5,
                values = "brewer.greens")) +
  tm_borders(fill_alpha = 0.5)
```

Reversing the colour scheme:

```{r}
tm_shape(mpsz_pop2020)+
  tm_polygons("DEPENDENCY",
              fill.scale = tm_scale_intervals(
                style = "quantile",
                n = 5,
                values = "-brewer.greens")) +
  tm_borders(fill_alpha = 0.5)
```

### 21.4.5 Map Layouts

#### 21.4.5.1 Map Legend

```{r}
tm_shape(mpsz_pop2020)+
  tm_polygons("DEPENDENCY",
              fill.scale = tm_scale_intervals(
                style = "jenks",
                n = 5,
                values = "brewer.greens"),
              fill.legend = tm_legend(
                title = "Dependency ratio")) +
  tm_borders(fill_alpha = 0.5) +
  tm_title("Distribution of Dependency Ratio by planning subzone \n(Jenks classification)")
```

#### 21.4.5.2 Map style

```{r}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY", 
          style = "quantile", 
          palette = "-Greens") +
  tm_borders(alpha = 0.5) +
  tmap_style("classic")
```

#### 21.4.5.3 Cartographic Furniture

```{r}
tm_shape(mpsz_pop2020)+
  tm_polygons(fill = "DEPENDENCY", 
              fill.scale = tm_scale_intervals(
                style = "quantile",
                n = 5,
                values = "brewer.blues"),
              fill.legend = tm_legend(
                title = "Dependency ratio")) +
  tm_title("Distribution of Dependency Ratio by planning subzone") +
  tm_layout(frame = TRUE) +
  tm_borders(fill_alpha = 0.5) +
  tm_compass(type="8star", size = 2) +
  tm_grid(alpha =0.2) +
  tm_credits("Source: Planning Sub-zone boundary from Urban Redevelopment Authorithy (URA)\n and Population data from Department of Statistics DOS", 
             position = c("left", "bottom"))
```

Reset to default style:

```{r}
tmap_style("white")
```

### 21.4.6 Drawing Small Multiple Choropleth Maps

#### 21.4.6.1 By assigning multiple values to at least one of the aesthetic arguments

```{r}
tm_shape(mpsz_pop2020)+
  tm_fill(c("YOUNG", "AGED"),
          style = "equal", 
          palette = "Blues") +
  tm_layout(legend.position = c("right", "bottom")) +
  tm_borders(alpha = 0.5) +
  tmap_style("white")
```

Multiple values with different styles:

```{r}
tm_shape(mpsz_pop2020)+ 
  tm_polygons(c("DEPENDENCY","AGED"),
              style = c("equal", "quantile"), 
              palette = list("Blues","Greens")) +
  tm_layout(legend.position = c("right", "bottom"))
```

#### 21.4.6.2 By defining a group-by variable in tm_facets()

```{r}
tm_shape(mpsz_pop2020) +
  tm_fill("DEPENDENCY",
          style = "quantile",
          palette = "Blues",
          thres.poly = 0) + 
  tm_facets(by="REGION_N", 
            free.coords=TRUE) +
  tm_layout(legend.show = FALSE,
            title.position = c("center", "center"), 
            title.size = 20) +
  tm_borders(alpha = 0.5)
```

#### 21.4.6.3 By creating multiple stand-alone maps with tmap_arrange()

```{r}
youngmap <- tm_shape(mpsz_pop2020)+ 
  tm_polygons("YOUNG", 
              style = "quantile", 
              palette = "Blues")

agedmap <- tm_shape(mpsz_pop2020)+ 
  tm_polygons("AGED", 
              style = "quantile", 
              palette = "Blues")

tmap_arrange(youngmap, agedmap, asp=1, ncol=2)
```

### 21.4.7 Mapping Spatial Object Meeting a Selection Criterion

```{r}
tm_shape(mpsz_pop2020[mpsz_pop2020$REGION_N=="CENTRAL REGION", ])+
  tm_fill("DEPENDENCY", 
          style = "quantile", 
          palette = "Blues", 
          legend.hist = TRUE, 
          legend.is.portrait = TRUE,
          legend.hist.z = 0.1) +
  tm_layout(legend.outside = TRUE,
            legend.height = 0.45, 
            legend.width = 5.0,
            legend.position = c("right", "bottom"),
            frame = FALSE) +
  tm_borders(alpha = 0.5)
```
