---
title: "Hands-on_Ex09c"
author: "Huang Zhanyi"
date: "June 18, 2026"
date-modified: "last-modified"
execute: 
  eval: true
  echo: true
  warning: false
  freeze: true
---

# 23 Analytical Mapping

## 23.1 Overview

### 23.1.1 Objectives

In this in-class exercise, you will gain hands-on experience on using appropriate R methods to plot analytical maps.

### 23.1.2 Learning Outcome

By the end of this in-class exercise, you will be able to use appropriate functions of tmap and tidyverse to perform the following tasks:

- Importing geospatial data in rds format into R environment.
- Creating cartographic quality choropleth maps by using appropriate tmap functions.
- Creating rate map
- Creating percentile map
- Creating boxmap

## 23.2 Getting Started

### 23.2.1 Installing and Loading Packages

Install and load sf, tmap and tidyverse packages into R environment:

```{r}
pacman::p_load(sf, tmap, tidyverse)
```

### 23.2.2 Importing Data

Import NGA_wp.rds into R environment:

```{r}
NGA_wp <- readRDS("data/NGA_wp.rds")
```

## 23.3 Basic Choropleth Mapping

### 23.3.1 Visualising Distribution of Non-functional Water Point

Plot a choropleth map showing the distribution of non-function water point by LGA:

```{r}
p2 <- tm_shape(NGA_wp) +
  tm_fill("wp_nonfunctional",
          n = 5,
          style = "jenks",
          title = "Non-functional\nwater points") +
  tm_borders(lwd = 0.1) +
  tm_layout(main.title = "Distribution of Non-functional Water Points by LGA",
            legend.position = c("left", "bottom"))

p1 <- tm_shape(NGA_wp) +
  tm_fill("total_wp",
          n = 5,
          style = "jenks",
          title = "Total\nwater points") +
  tm_borders(lwd = 0.1) +
  tm_layout(main.title = "Distribution of Total Water Points by LGA",
            legend.position = c("left", "bottom"))

tmap_arrange(p2, p1, nrow = 1)
```

## 23.4 Choropleth Map for Rates

In much of our readings we have now seen the importance to map rates rather than counts of things, and that is for the simple reason that water points are not equally distributed in space.

### 23.4.1 Deriving Proportion of Functional Water Points and Non-Functional Water Points

Derive proportion of functional and non-functional water points:

```{r}
NGA_wp <- NGA_wp %>%
  mutate(pct_functional = wp_functional/total_wp) %>%
  mutate(pct_nonfunctional = wp_nonfunctional/total_wp)
```

### 23.4.2 Plotting Map of Rate

Plot a choropleth map showing the distribution of percentage functional water point by LGA:

```{r}
tm_shape(NGA_wp) +
  tm_fill("pct_functional",
          n = 5,
          style = "pretty",
          palette = "Blues",
          legend.hist = TRUE,
          title = "Percentage\nfunctional") +
  tm_borders(lwd = 0.1) +
  tm_layout(main.title = "Distribution of Functional Water Points Percentage by LGA",
            legend.position = c("right", "bottom"))
```

## 23.5 Extreme Value Maps

Extreme value maps are variations of common choropleth maps where the classification is designed to highlight extreme values at the lower and upper end of the scale, with the goal of identifying outliers.

### 23.5.1 Percentile Map

The percentile map is a special type of quantile map with six specific categories: 0-1%, 1-10%, 10-50%, 50-90%, 90-99%, and 99-100%.

#### 23.5.1.1 Data Preparation

Step 1: Exclude records with NA:

```{r}
NGA_wp <- NGA_wp %>%
  drop_na()
```

Step 2: Creating customised classification and extracting values:

```{r}
percent <- c(0,.01,.1,.5,.9,.99,1)
var <- NGA_wp["pct_functional"] %>%
  st_set_geometry(NULL)
quantile(var[,1], percent)
```

#### 23.5.1.3 Creating the get.var Function

Function to extract a variable as a vector out of an sf data frame:

```{r}
get.var <- function(vname,df) {
  v <- df[vname] %>% 
    st_set_geometry(NULL)
  v <- unname(v[,1])
  return(v)
}
```

#### 23.5.1.4 A Percentile Mapping Function

Function to create a percentile map:

```{r}
percentmap <- function(vnam, df, legtitle=NA, mtitle="Percentile Map"){
  percent <- c(0,.01,.1,.5,.9,.99,1)
  var <- get.var(vnam, df)
  bperc <- quantile(var, percent)
  tm_shape(df) +
  tm_polygons() +
  tm_shape(df) +
     tm_polygons(vnam,
             title=legtitle,
             breaks=bperc,
             palette="Blues",
             labels=c("< 1%", "1% - 10%", "10% - 50%", "50% - 90%", "90% - 99%", "> 99%"))  +
  tm_borders() +
  tm_layout(main.title = mtitle, 
            title.position = c("right","bottom"))
}
```

#### 23.5.1.5 Test Drive the Percentile Mapping Function

Run the percentile mapping function:

```{r}
percentmap("total_wp", NGA_wp)
```

### 23.5.2 Box Map

A box map is an augmented quartile map, with an additional lower and upper category. Displaying summary statistics on a choropleth map by using the basic principles of boxplot.

#### 23.5.2.1 Creating the boxbreaks Function

Function to create break points for a box map:

```{r}
boxbreaks <- function(v,mult=1.5) {
  qv <- unname(quantile(v))
  iqr <- qv[4] - qv[2]
  upfence <- qv[4] + mult * iqr
  lofence <- qv[2] - mult * iqr
  # initialize break points vector
  bb <- vector(mode="numeric",length=7)
  # logic for lower and upper fences
  if (lofence < qv[1]) {  # no lower outliers
    bb[1] <- lofence
    bb[2] <- floor(qv[1])
  } else {
    bb[2] <- lofence
    bb[1] <- qv[1]
  }
  if (upfence > qv[5]) { # no upper outliers
    bb[7] <- upfence
    bb[6] <- ceiling(qv[5])
  } else {
    bb[6] <- upfence
    bb[7] <- qv[5]
  }
  bb[3:5] <- qv[2:4]
  return(bb)
}
```

#### 23.5.2.2 Creating the get.var Function

Function to extract a variable as a vector out of an sf data frame:

```{r}
get.var <- function(vname,df) {
  v <- df[vname] %>% st_set_geometry(NULL)
  v <- unname(v[,1])
  return(v)
}
```

#### 23.5.2.3 Test Drive the Newly Created Function

Test the boxbreaks function:

```{r}
var <- get.var("wp_nonfunctional", NGA_wp) 
boxbreaks(var)
```

#### 23.5.2.4 Boxmap Function

Function to create a box map:

```{r}
boxmap <- function(vnam, df, 
                   legtitle=NA,
                   mtitle="Box Map",
                   mult=1.5){
  var <- get.var(vnam,df)
  bb <- boxbreaks(var)
  tm_shape(df) +
    tm_polygons() +
  tm_shape(df) +
     tm_fill(vnam,title=legtitle,
             breaks=bb,
             palette="Blues",
             labels = c("lower outlier", 
                        "< 25%", 
                        "25% - 50%", 
                        "50% - 75%",
                        "> 75%", 
                        "upper outlier"))  +
  tm_borders() +
  tm_layout(main.title = mtitle, 
            title.position = c("left",
                               "top"))
}
```

Test the boxmap function:

```{r}
tmap_mode("plot")
boxmap("wp_nonfunctional", NGA_wp)
```
