---
title: "Hands-on_Ex08"
author: "Huang Zhanyi"
date: "June 10, 2026"
date-modified: "last-modified"
execute: 
  eval: true
  echo: true
  warning: false
  freeze: true
---

## Learning Outcome

By the end of this hands-on exercise you will be able to create the following data visualisations using R packages:

- Plotting a calendar heatmap by using `ggplot2` functions
- Plotting a cycle plot by using `ggplot2` functions
- Plotting a slopegraph
- Plotting a horizon chart

## Getting Started

### Loading Required Packages

The following R packages are required for this exercise. The code chunk below checks, installs (if not yet installed), and loads them into the R environment.

```{r}
pacman::p_load(scales, viridis, lubridate, ggthemes,
               gridExtra, readxl, knitr, data.table,
               tidyverse)

eval(parse(text = readLines("https://raw.githubusercontent.com/ibecav/CGPfunctions/master/R/NewggslopegraphY.R")))
```

## Plotting Calendar Heatmap

In this section, you will learn how to plot a calendar heatmap programmatically by using the `ggplot2` package.

By the end of this section, you will be able to:

- Plot a calendar heatmap by using `ggplot2` functions and extensions
- Write functions using R programming
- Derive specific date and time related fields using base R and `lubridate` packages
- Perform data preparation tasks using `tidyr` and `dplyr` packages

### The Data

For the purpose of this hands-on exercise, `eventlog.csv` will be used. This data file consists of 199,999 rows of time-series cyber attack records by country.

### Importing the Data

```{r}
attacks <- read_csv("/Users/zhanyi/Desktop/M Y1S2/ISSS608/ISSS608-VAA/project 1/ISSS608/Hands-on_Ex/Hands-on_Ex08/data/eventlog.csv")
```

### Examining the Data Structure

It is always good practice to examine the imported data frame before further analysis is performed. `kable()` is used to review the structure of the imported data frame.

```{r}
kable(head(attacks))
```

There are three columns: `timestamp`, `source_country`, and `tz`.

- `timestamp` stores date-time values in POSIXct format.
- `source_country` stores the source of the attack in ISO 3166-1 alpha-2 country code.
- `tz` stores the time zone of the source IP address.

### Data Preparation

**Step 1: Deriving weekday and hour of day fields**

Before plotting the calendar heatmap, two new fields — `wkday` and `hour` — need to be derived. A function is written to perform this task.

::: callout-note
`ymd_hms()` and `hour()` are from the `lubridate` package, and `weekdays()` is a base R function.
:::

```{r}
make_hr_wkday <- function(ts, sc, tz) {
  real_times <- ymd_hms(ts,
                        tz = tz[1],
                        quiet = TRUE)
  dt <- data.table(source_country = sc,
                   wkday = weekdays(real_times),
                   hour = hour(real_times))
  return(dt)
}
```

**Step 2: Deriving the attacks tibble data frame**

::: callout-note
Beside extracting the necessary data into the `attacks` data frame, `mutate()` of `dplyr` is used to convert `wkday` and `hour` fields into factors so they will be ordered when plotting.
:::

```{r}
wkday_levels <- c('Saturday', 'Friday',
                  'Thursday', 'Wednesday',
                  'Tuesday', 'Monday',
                  'Sunday')

attacks <- attacks %>%
  group_by(tz) %>%
  do(make_hr_wkday(.$timestamp,
                   .$source_country,
                   .$tz)) %>%
  ungroup() %>%
  mutate(wkday = factor(
    wkday, levels = wkday_levels),
    hour = factor(
      hour, levels = 0:23))
```

The table below shows the tidy tibble after processing:

```{r}
kable(head(attacks))
```

### Building the Calendar Heatmap

```{r}
grouped <- attacks %>%
  count(wkday, hour) %>%
  ungroup() %>%
  na.omit()

ggplot(grouped,
       aes(hour,
           wkday,
           fill = n)) +
  geom_tile(color = "white",
            size = 0.1) +
  theme_tufte(base_family = "Helvetica") +
  coord_equal() +
  scale_fill_gradient(name = "# of attacks",
                      low = "sky blue",
                      high = "dark blue") +
  labs(x = NULL,
       y = NULL,
       title = "Attacks by weekday and time of day") +
  theme(axis.ticks = element_blank(),
        plot.title = element_text(hjust = 0.5),
        legend.title = element_text(size = 8),
        legend.text = element_text(size = 6))
```

::: callout-note
Things to learn from the code chunk:

- A tibble data table called `grouped` is derived by aggregating attacks by `wkday` and `hour` fields.
- A new field `n` is derived using `group_by()` and `count()`.
- `na.omit()` is used to exclude missing values.
- `geom_tile()` plots tiles (grids) at each x and y position. The `color` and `size` arguments specify the border colour and line size.
- `theme_tufte()` of `ggthemes` removes unnecessary chart junk.
- `coord_equal()` ensures an aspect ratio of 1:1.
- `scale_fill_gradient()` creates a two-colour gradient (low–high).
:::

### Building Multiple Calendar Heatmaps

**Challenge:** Build multiple heatmaps for the top four countries with the highest number of attacks.

**Step 1: Deriving attack by country object**

To identify the top 4 countries with the highest number of attacks:

```{r}
attacks_by_country <- count(attacks, source_country) %>%
  mutate(percent = percent(n / sum(n))) %>%
  arrange(desc(n))
```

**Step 2: Preparing the tidy data frame**

Extract attack records of the top 4 countries and save in a new tibble `top4_attacks`.

```{r}
top4 <- attacks_by_country$source_country[1:4]

top4_attacks <- attacks %>%
  filter(source_country %in% top4) %>%
  count(source_country, wkday, hour) %>%
  ungroup() %>%
  mutate(source_country = factor(
    source_country, levels = top4)) %>%
  na.omit()
```

**Step 3: Plotting the Multiple Calendar Heatmap**

```{r}
ggplot(top4_attacks,
       aes(hour,
           wkday,
           fill = n)) +
  geom_tile(color = "white",
            size = 0.1) +
  theme_tufte(base_family = "Helvetica") +
  coord_equal() +
  scale_fill_gradient(name = "# of attacks",
                      low = "sky blue",
                      high = "dark blue") +
  facet_wrap(~source_country, ncol = 2) +
  labs(x = NULL, y = NULL,
       title = "Attacks on top 4 countries by weekday and time of day") +
  theme(axis.ticks = element_blank(),
        axis.text.x = element_text(size = 7),
        plot.title = element_text(hjust = 0.5),
        legend.title = element_text(size = 8),
        legend.text = element_text(size = 6))
```

## Plotting Cycle Plot

In this section, you will learn how to plot a cycle plot showing the time-series patterns and trend of visitor arrivals from Vietnam using `ggplot2`.

### Step 1: Data Import

`arrivals_by_air.xlsx` is imported using `read_excel()` of the `readxl` package.

```{r}
air <- read_excel("data/arrivals_by_air.xlsx")
```

### Step 2: Deriving Month and Year Fields

Two new fields — `month` and `year` — are derived from the `Month-Year` field.

```{r}
air$month <- factor(month(air$`Month-Year`),
                    levels = 1:12,
                    labels = month.abb,
                    ordered = TRUE)
air$year <- year(ymd(air$`Month-Year`))
```

### Step 3: Extracting the Target Country

Extract data for Vietnam only.

```{r}
Vietnam <- air %>%
  select(`Vietnam`, month, year) %>%
  filter(year >= 2010)
```

### Step 4: Computing Year Average Arrivals by Month

`group_by()` and `summarise()` of `dplyr` are used to compute year average arrivals by month.

```{r}
hline.data <- Vietnam %>%
  group_by(month) %>%
  summarise(avgvalue = mean(`Vietnam`))
```

### Step 5: Plotting the Cycle Plot

```{r}
ggplot() +
  geom_line(data = Vietnam,
            aes(x = year,
                y = `Vietnam`,
                group = month),
            colour = "black") +
  geom_hline(aes(yintercept = avgvalue),
             data = hline.data,
             linetype = 6,
             colour = "red",
             size = 0.5) +
  facet_grid(~month) +
  labs(axis.text.x = element_blank(),
       title = "Visitor arrivals from Vietnam by air, Jan 2010-Dec 2019") +
  xlab("") +
  ylab("No. of Visitors") +
  theme_tufte(base_family = "Helvetica")
```

## Plotting Slopegraph

In this section you will learn how to plot a slopegraph using R.

::: callout-note
Before getting started, make sure that `CGPfunctions` has been installed and loaded. Refer to [Using newggslopegraph](https://cran.r-project.org/web/packages/CGPfunctions/vignettes/Using-newggslopegraph.html) to learn more about the function.
:::

### Step 1: Data Import

```{r}
rice <- read_csv("data/rice.csv")
```

### Step 2: Plotting the Slopegraph

```{r}
rice %>%
  mutate(Year = factor(Year)) %>%
  filter(Year %in% c(1961, 1980)) %>%
  ggplot(aes(x = Year, y = Yield, group = Country, colour = Country)) +
  geom_line(linewidth = 1) +
  geom_point(size = 3) +
  geom_text(data = . %>% filter(Year == "1961"),
            aes(label = paste0(Country, " ", round(Yield, 0))),
            hjust = 1.15, size = 3) +
  geom_text(data = . %>% filter(Year == "1980"),
            aes(label = paste0(Country, " ", round(Yield, 0))),
            hjust = -0.15, size = 3) +
  theme_tufte() +
  theme(legend.position = "none",
        plot.margin = margin(l = 80, r = 80)) +
  labs(title = "Rice Yield of Top 11 Asian Counties",
       subtitle = "1961-1980",
       caption = "Prepared by: Dr. Kam Tin Seong",
       x = NULL, y = "Yield")
```
