---
title: "Hands-on_Ex07a"
author: "Huang Zhanyi"
date: "June 4, 2026"
date-modified: "last-modified"
execute: 
  eval: true
  echo: true
  warning: false
  freeze: true
---

## 29.2 Getting Started

### 29.2.1 Installing and Launching R Packages

```{r}
pacman::p_load(tidytext, widyr, wordcloud, DT, ggwordcloud, textplot, lubridate, hms,
tidyverse, tidygraph, ggraph, igraph)
```

------------------------------------------------------------------------

## 29.3 Importing Multiple Text Files from Multiple Folders

### 29.3.1 Creating a Folder List

```{r}
news20 <- "/Users/zhanyi/Desktop/M Y1S2/ISSS608/ISSS608-VAA/project 1/ISSS608/Hands-on_Ex/Hands-on_Ex07b/data/20news/"
```

### 29.3.2 Define a Function to Read All Files from a Folder into a Data Frame

```{r}
read_folder <- function(infolder) {
  tibble(file = dir(infolder, 
                    full.names = TRUE)) %>%
    mutate(text = map(file, 
                      read_lines)) %>%
    transmute(id = basename(file), 
              text) %>%
    unnest(text)
}
```

------------------------------------------------------------------------

## 29.4 Importing Multiple Text Files from Multiple Folders

### 29.4.1 Reading in All the Messages from the 20news Folder

```{r}
raw_text <- tibble(folder = 
                     dir(news20, 
                         full.names = TRUE)) %>%
  mutate(folder_out = map(folder, 
                          read_folder)) %>%
  unnest(cols = c(folder_out)) %>%
  transmute(newsgroup = basename(folder), 
            id, text)
write_rds(raw_text, "data/rds/news20.rds")
```

------------------------------------------------------------------------

## 29.5 Initial EDA

Visualise frequency of messages by newsgroup.

```{r}
raw_text %>%
  group_by(newsgroup) %>%
  summarize(messages = n_distinct(id)) %>%
  ggplot(aes(x = messages, y = reorder(newsgroup, messages))) +
  geom_col(fill = "lightblue") +
  geom_text(aes(label = messages), hjust = -0.3, size = 3) +
  scale_x_continuous(limits = c(0, 13)) +
  labs(x = "messages", y = NULL) +
  theme_minimal() +
  theme(axis.text.y = element_text(size = 9))
```

------------------------------------------------------------------------

## 29.6 Introducing tidytext

### 29.6.1 Removing Header and Automated Email Signatures

Remove email headers and signatures (lines after `--`).

```{r}
cleaned_text <- raw_text %>%
  group_by(newsgroup, id) %>%
  filter(cumsum(text == "") > 0,
         cumsum(str_detect(
           text, "^--")) == 0) %>%
  ungroup()
```

### 29.6.2 Removing Lines with Nested Text Representing Quotes from Other Users

```{r}
cleaned_text <- cleaned_text %>%
  filter(str_detect(text, "^[^>]+[A-Za-z\\d]")
         | text == "",
         !str_detect(text, 
                     "writes(:|\\.\\.\\.)$"),
         !str_detect(text, 
                     "^In article <")
  )
```

### 29.6.3 Text Data Processing

Tokenise words and remove stop words.

```{r}
usenet_words <- cleaned_text %>%
  unnest_tokens(word, text) %>%
  filter(str_detect(word, "[a-z']$"),
         !word %in% stop_words$word)
```

Count the most common words across all newsgroups.

```{r}
usenet_words %>%
  count(word, sort = TRUE)
```

Count words within each newsgroup.

```{r}
words_by_newsgroup <- usenet_words %>%
  count(newsgroup, word, sort = TRUE) %>%
  ungroup()
```

### 29.6.4 Visualising Words in Newsgroups

Plot a static word cloud using `wordcloud`.

```{r}
wordcloud(words_by_newsgroup$word,
          words_by_newsgroup$n,
          max.words = 300)
```

### 29.6.5 Visualising Words in Newsgroups with ggwordcloud

Plot faceted word clouds per newsgroup using `ggwordcloud`.

```{r}
set.seed(1234)
words_by_newsgroup %>%
  filter(n > 0) %>%
  ggplot(aes(label = word, size = n)) +
  geom_text_wordcloud() +
  theme_minimal() +
  facet_wrap(~newsgroup) +
  scale_size_area(max_size = 8)
```

------------------------------------------------------------------------

## 29.7 Basic Concept of TF-IDF

### 29.7.1 Computing tf-idf Within Newsgroups

```{r}
tf_idf <- words_by_newsgroup %>%
  bind_tf_idf(word, newsgroup, n) %>%
  arrange(desc(tf_idf))
```

### 29.7.3 Visualising tf-idf as Interactive Table

```{r}
DT::datatable(tf_idf, filter = 'top') %>% 
  formatRound(columns = c('tf', 'idf', 
                          'tf_idf'), 
              digits = 3) %>%
  formatStyle(0, 
              target = 'row', 
              lineHeight='25%')
```

### 29.7.4 Visualising tf-idf Within Newsgroups

Facet bar charts of tf-idf values for science-related newsgroups.

```{r}
tf_idf %>%
  filter(str_detect(newsgroup, "^sci\\.")) %>%
  group_by(newsgroup) %>%
  slice_max(tf_idf, 
            n = 12) %>%
  ungroup() %>%
  mutate(word = reorder(word, 
                        tf_idf)) %>%
  ggplot(aes(tf_idf, 
             word, 
             fill = newsgroup)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~ newsgroup, 
             scales = "free") +
  labs(x = "tf-idf", 
       y = NULL)
```

### 29.7.5 Counting and Correlating Pairs of Words with the widyr Package

Compute pairwise correlation between newsgroups based on shared words.

```{r}
newsgroup_cors <- words_by_newsgroup %>%
  pairwise_cor(newsgroup, 
               word, 
               n, 
               sort = TRUE)
```

### 29.7.6 Visualising Correlation as a Network

```{r}
set.seed(2017)

newsgroup_cors %>%
  filter(correlation > .025) %>%
  graph_from_data_frame() %>%
  ggraph(layout = "fr") +
  geom_edge_link(aes(alpha = correlation, 
                     width = correlation)) +
  geom_node_point(size = 6, 
                  color = "lightblue") +
  geom_node_text(aes(label = name),
                 color = "red",
                 repel = TRUE) +
  theme_void()
```

### 29.7.7 Bigram

Create a bigram data frame using `unnest_tokens()`.

```{r}
bigrams <- cleaned_text %>%
  unnest_tokens(bigram, 
                text, 
                token = "ngrams", 
                n = 2)

bigrams
```

### 29.7.8 Counting Bigrams

```{r}
bigrams_count <- bigrams %>%
  filter(bigram != 'NA') %>%
  count(bigram, sort = TRUE)

bigrams_count
```

### 29.7.9 Cleaning Bigram

Separate bigrams into two words and remove stop words.

```{r}
bigrams_separated <- bigrams %>%
  filter(bigram != 'NA') %>%
  separate(bigram, c("word1", "word2"), 
           sep = " ")

bigrams_filtered <- bigrams_separated %>%
  filter(!word1 %in% stop_words$word) %>%
  filter(!word2 %in% stop_words$word)

bigrams_filtered
```

### 29.7.10 Counting the Bigram Again

```{r}
bigram_counts <- bigrams_filtered %>% 
  count(word1, word2, sort = TRUE)
```

### 29.7.11 Create a Network Graph from Bigram Data Frame

```{r}
bigram_graph <- bigram_counts %>%
  filter(n > 3) %>%
  graph_from_data_frame()
bigram_graph
```

### 29.7.12 Visualising a Network of Bigrams with ggraph

```{r}
set.seed(1234)

ggraph(bigram_graph, layout = "fr") +
  geom_edge_link() +
  geom_node_point() +
  geom_node_text(aes(label = name), 
                 vjust = 1, 
                 hjust = 1)
```

### 29.7.13 Revised Version

Add edge transparency, directional arrows, and improved styling.

```{r}
set.seed(1234)

a <- grid::arrow(type = "closed", 
                 length = unit(.15,
                               "inches"))

ggraph(bigram_graph, 
       layout = "fr") +
  geom_edge_link(aes(edge_alpha = n), 
                 show.legend = FALSE,
                 arrow = a, 
                 end_cap = circle(.07,
                                  'inches')) +
  geom_node_point(color = "lightblue", 
                  size = 5) +
  geom_node_text(aes(label = name), 
                 vjust = 1, 
                 hjust = 1) +
  theme_void()
```
