---
title: "Hands-on_Ex07a"
author: "Huang Zhanyi"
date: "June 4, 2026"
date-modified: "last-modified"
execute: 
  eval: true
  echo: true
  warning: false
  freeze: true
---

## 27.1 Overview

This exercise covers modelling, analysing, and visualising network data using R — including graph objects, network metrics, and interactive graphs.

------------------------------------------------------------------------

## 27.2 Getting Started

### 27.2.1 Installing and Launching R Packages

Install and load the required packages.

```{r}
pacman::p_load(igraph, tidygraph, ggraph, 
               visNetwork, lubridate, clock,
               tidyverse, graphlayouts, 
               concaveman, ggforce)
```

------------------------------------------------------------------------

## 27.3 The Data

Two datasets from an oil exploration company:

- **Edges**: `GAStech-email_edges.csv` — 9,063 email records between 55 employees over 2 weeks
- **Nodes**: `GAStech_email_nodes.csv` — names, departments, and titles of 55 employees

### 27.3.3 Importing Network Data

```{r}
GAStech_nodes <- read_csv("/Users/zhanyi/Desktop/M Y1S2/ISSS608/ISSS608-VAA/project 1/ISSS608/Hands-on_Ex/Hands-on_Ex07a/data/GAStech_email_node.csv")
GAStech_edges <- read_csv("/Users/zhanyi/Desktop/M Y1S2/ISSS608/ISSS608-VAA/project 1/ISSS608/Hands-on_Ex/Hands-on_Ex07a/data/edgev2.csv")
```

### 27.3.4 Reviewing the Imported Data

```{r}
glimpse(GAStech_edges)
```

### 27.3.5 Wrangling Time

Convert `SentDate` from character to Date and extract the weekday.

```{r}
GAStech_edges <- GAStech_edges %>%
  mutate(SendDate = dmy(SentDate),
         Weekday = wday(SendDate,
                        label = TRUE,
                        abbr = FALSE))
```

### 27.3.7 Wrangling Attributes

Aggregate individual email records by sender, receiver, and weekday. A `Weight` column counts emails per group.

```{r}
GAStech_edges_aggregated <- GAStech_edges %>%
  filter(MainSubject == "Work related") %>%
  group_by(source, target, Weekday) %>%
    summarise(Weight = n()) %>%
  filter(source!=target) %>%
  filter(Weight > 1) %>%
  ungroup()
```

------------------------------------------------------------------------

## 27.4 Creating Network Objects Using tidygraph

### 27.4.3 Using tbl_graph() to Build the tidygraph Data Model

```{r}
GAStech_graph <- tbl_graph(nodes = GAStech_nodes,
                           edges = GAStech_edges_aggregated, 
                           directed = TRUE)
```

### 27.4.6 Changing the Active Object

Switch to the edges tibble and reorder by descending weight.

```{r}
GAStech_graph %>%
  activate(edges) %>%
  arrange(desc(Weight))
```

------------------------------------------------------------------------

## 27.5 Plotting Static Network Graphs with ggraph

### 27.5.1 Plotting a Basic Network Graph

```{r}
ggraph(GAStech_graph) +
  geom_edge_link() +
  geom_node_point()
```

### 27.5.2 Changing the Default Network Graph Theme

Removes axes, grids, and borders for a cleaner look.

```{r}
g <- ggraph(GAStech_graph) + 
  geom_edge_link(aes()) +
  geom_node_point(aes())

g + theme_graph()
```

### 27.5.3 Changing the Colouring of the Plot

```{r}
g <- ggraph(GAStech_graph) + 
  geom_edge_link(aes(colour = 'grey50')) +
  geom_node_point(aes(colour = 'grey40'))

g + theme_graph(background = 'grey10',
                text_colour = 'white')
```

### 27.5.5 Fruchterman and Reingold Layout

```{r}
g <- ggraph(GAStech_graph, 
            layout = "fr") +
  geom_edge_link(aes()) +
  geom_node_point(aes())

g + theme_graph()
```

### 27.5.6 Modifying Network Nodes

Colour nodes by department.

```{r}
g <- ggraph(GAStech_graph, 
            layout = "nicely") + 
  geom_edge_link(aes()) +
  geom_node_point(aes(colour = Department, 
                      size = 3))

g + theme_graph()
```

### 27.5.7 Modifying Edges

Map edge width to `Weight` and add transparency.

```{r}
g <- ggraph(GAStech_graph, 
            layout = "nicely") +
  geom_edge_link(aes(width=Weight), 
                 alpha=0.2) +
  scale_edge_width(range = c(0.1, 5)) +
  geom_node_point(aes(colour = Department), 
                  size = 3)

g + theme_graph()
```

------------------------------------------------------------------------

## 27.6 Creating Facet Graphs

### 27.6.1 Working with facet_edges()

Split the graph into panels by `Weekday`.

```{r}
set_graph_style()

g <- ggraph(GAStech_graph, 
            layout = "nicely") + 
  geom_edge_link(aes(width=Weight), 
                 alpha=0.2) +
  scale_edge_width(range = c(0.1, 5)) +
  geom_node_point(aes(colour = Department), 
                  size = 2)

g + facet_edges(~Weekday)
```

### 27.6.2 Moving the Legend

```{r}
set_graph_style()

g <- ggraph(GAStech_graph, 
            layout = "nicely") + 
  geom_edge_link(aes(width=Weight), 
                 alpha=0.2) +
  scale_edge_width(range = c(0.1, 5)) +
  geom_node_point(aes(colour = Department), 
                  size = 2) +
  theme(legend.position = 'bottom')
  
g + facet_edges(~Weekday)
```

### 27.6.3 A Framed Facet Graph

```{r}
set_graph_style() 

g <- ggraph(GAStech_graph, 
            layout = "nicely") + 
  geom_edge_link(aes(width=Weight), 
                 alpha=0.2) +
  scale_edge_width(range = c(0.1, 5)) +
  geom_node_point(aes(colour = Department), 
                  size = 2)
  
g + facet_edges(~Weekday) +
  th_foreground(foreground = "grey80",  
                border = TRUE) +
  theme(legend.position = 'bottom')
```

### 27.6.4 Working with facet_nodes()

```{r}
set_graph_style()

g <- ggraph(GAStech_graph, 
            layout = "nicely") + 
  geom_edge_link(aes(width=Weight), 
                 alpha=0.2) +
  scale_edge_width(range = c(0.1, 5)) +
  geom_node_point(aes(colour = Department), 
                  size = 2)
  
g + facet_nodes(~Department)+
  th_foreground(foreground = "grey80",  
                border = TRUE) +
  theme(legend.position = 'bottom')
```

------------------------------------------------------------------------

## 27.7 Network Metrics Analysis

### 27.7.1 Computing Centrality Indices

Calculate betweenness centrality and map it to node size.

```{r}
g <- GAStech_graph %>%
  mutate(betweenness_centrality = centrality_betweenness()) %>%
  ggraph(layout = "fr") + 
  geom_edge_link(aes(width=Weight), 
                 alpha=0.2) +
  scale_edge_width(range = c(0.1, 5)) +
  geom_node_point(aes(colour = Department,
            size=betweenness_centrality))
g + theme_graph()
```

### 27.7.2 Visualising Network Metrics

From ggraph v2.0, centrality measures can be computed directly inside the `ggraph` call without precomputing.

```{r}
g <- GAStech_graph %>%
  ggraph(layout = "fr") + 
  geom_edge_link(aes(width=Weight), 
                 alpha=0.2) +
  scale_edge_width(range = c(0.1, 5)) +
  geom_node_point(aes(colour = Department, 
                      size = centrality_betweenness()))
g + theme_graph()
```

### 27.7.3 Visualising Community

Detect communities using edge-betweenness and colour nodes by community.

```{r}
g <- GAStech_graph %>%
  mutate(community = as.factor(
    group_edge_betweenness(
      weights = Weight, 
      directed = TRUE))) %>%
  ggraph(layout = "fr") + 
  geom_edge_link(
    aes(
      width=Weight), 
    alpha=0.2) +
  scale_edge_width(
    range = c(0.1, 5)) +
  geom_node_point(
    aes(colour = community))  

g + theme_graph()
```

Use `geom_mark_hull()` from `ggforce` to draw convex hulls around communities.

```{r}
g <- GAStech_graph %>%
  activate(nodes) %>%
  mutate(community = as.factor(
    group_optimal(weights = Weight)),
         betweenness_measure = centrality_betweenness()) %>%
  ggraph(layout = "fr") +
  geom_mark_hull(
    aes(x, y, 
        group = community, 
        fill = community),  
    alpha = 0.2,  
    expand = unit(0.3, "cm"),  
    radius = unit(0.3, "cm")  
  ) + 
  geom_edge_link(aes(width=Weight), 
                 alpha=0.2) +
  scale_edge_width(range = c(0.1, 5)) +
  geom_node_point(aes(fill = Department,
                      size = betweenness_measure),
                      color = "black",
                      shape = 21)
  
g + theme_graph()
```

------------------------------------------------------------------------

## 27.8 Building Interactive Network Graph with visNetwork

### 27.8.1 Data Preparation

Re-aggregate edges using `left_join` to get proper `from`/`to` node IDs required by `visNetwork`.

```{r}
GAStech_edges_aggregated <- GAStech_edges %>%
  left_join(GAStech_nodes, by = c("sourceLabel" = "label")) %>%
  rename(from = id) %>%
  left_join(GAStech_nodes, by = c("targetLabel" = "label")) %>%
  rename(to = id) %>%
  filter(MainSubject == "Work related") %>%
  group_by(from, to) %>%
    summarise(weight = n()) %>%
  filter(from!=to) %>%
  filter(weight > 1) %>%
  ungroup()
```

### 27.8.2 Plotting the First Interactive Network Graph

```{r}
visNetwork(GAStech_nodes, 
           GAStech_edges_aggregated)
```

### 27.8.3 Working with Layout

```{r}
visNetwork(GAStech_nodes,
           GAStech_edges_aggregated) %>%
  visIgraphLayout(layout = "layout_with_fr") 
```

### 27.8.4 Working with Visual Attributes — Nodes

```{r}
visNetwork(GAStech_nodes,
           GAStech_edges_aggregated) %>%
  visIgraphLayout(layout = "layout_with_fr") %>%
  visLegend() %>%
  visLayout(randomSeed = 123)
```

### 27.8.5 Working with Visual Attributes — Edges

Add directional arrows and smooth curved edges.

```{r}
visNetwork(GAStech_nodes,
           GAStech_edges_aggregated) %>%
  visIgraphLayout(layout = "layout_with_fr") %>%
  visEdges(arrows = "to", 
           smooth = list(enabled = TRUE, 
                         type = "curvedCW")) %>%
  visLegend() %>%
  visLayout(randomSeed = 123)
```

### 27.8.6 Interactivity

Highlight nearest neighbours on click and add a node selection dropdown.

```{r}
visNetwork(GAStech_nodes,
           GAStech_edges_aggregated) %>%
  visIgraphLayout(layout = "layout_with_fr") %>%
  visOptions(highlightNearest = TRUE,
             nodesIdSelection = TRUE) %>%
  visLegend() %>%
  visLayout(randomSeed = 123)
```
